%% document class
\documentclass[10pt,oneside,leqno,openright]{memoir}

\input{macros}
\newcommand*{\sto}{\rightsquigarrow}


\begin{document}

\frontmatter
\thispagestyle{empty}
\begin{center}
  \LARGE
  {\bfseries Synthetic Homotopy Theory}\\[1cm]
  \large Exercises and Solutions for the\\[5mm]
  \Large Midlands Graduate School (MGS)\\
  \large 13--17 April 2026, Nottingham, UK\\[1cm]
  by\\
  Axel Ljungström and Aref Mohammadzadeh\\
  School of Computer Science\\
  University of Nottingham\\[1cm]
  Heavily based on notes from MGS 2024 by
  Ulrik Buchholtz and Mark Williams\\
  \Large NB\enspace Subject to change: Refresh if needed
\end{center}

\vfill

%\begin{abstract}
%  These lecture notes aim to introduce the field of synthetic homotopy theory to a (mainly) computer science audience. That is, we shall assume that the reader cares about types,
%\end{abstract}

\clearpage

\tableofcontents*

\mainmatter

%% \setcounter{chapter}{-1}

%% \chapter{Introduction}

\chapter{Truncated and connected types}
%\textcite{Shulman-logic-of-space}


\section{Exercises}
\begin{enumerate}
    \item Show that for $n \geq m$ and any type $A$ we have
    \[ \Trunc{\Trunc{A}_n}_m \simeq \Trunc{A}_m. \]

  \item\label{ex:n-conn-orth}
    Show that a type $X$ is $n$-connected if and only if, for every $n$-type $Y$,
    the diagonal map
    \[ \delta: Y \to (X \to Y), \quad y\mapsto_\delta (x\mapsto y),
    \]
    is an equivalence.

  \item Show that the map $\trunc\blank_n : X \to \Trunc X_n$ is $n$-connected.

    \item Consider a map $f : X \to Y$.
      We define its \textbf{$n$-image} to be the type $\im_n(f) :=
      (y : Y) \times \Trunc{\fib_f(y)}_n$.
    \begin{enumerate}
    \item Show that we can factorise $f$ through the $n$-image, that is,
      construct a commuting diagram:
        \[\begin{tikzcd}
            X \ar[rr, "f"] \ar[dr, swap, "p"] & & Y \\
            & \im_n(f) \ar[ur, swap, "i"]
        \end{tikzcd}\]
    \item Show that $p$ is $n$-connected and $i$ is $n$-truncated.
    \item Show that if $Z$ is any type, $g : X \to Z$ is $n$-connected, and $h:Z\to Y$ is $n$-truncated with $f = h\circ g$, then there is a unique equivalence $e : \im_n(f) \to Z$ with $e\circ p = g$ and $h\circ e = i$:
      \[
        \begin{tikzcd}
      & Z \ar[dr,"h"] & \\
      X \ar[rr,pos=0.7,"f"]\ar[dr,"p"'] \ar[ur,"g"]& & Y\\
      & \im_n(f)\ar[ur,"i"']\ar[uu,dotted,pos=0.3, "e"] &
    \end{tikzcd}
  \]
  (\emph{Hint})\enspace It's easier if you assume $g$ and $h$ are projections of
  dependent sums.
    \end{enumerate}

    \item (Eckmann--Hilton). All paths will take place in a fixed type $X$.
    \begin{enumerate}
        \item Given paths $p_0, p_1 : x = y$ and $q_0, q_1: y = z$, and higher paths $r : p_0 = p_1$ and $s : q_0 = q_1$, define their \textbf{horizontal multiplication} $s \ast r : q_0 \cdot p_0 = q_1 \cdot p_1$ by path induction on $r$ and $p_0$.
        \[\begin{tikzcd}
    	x & y & z
    	\arrow[""{name=0, anchor=center, inner sep=0}, "{p_1}"', bend right=40, from=1-1, to=1-2]
    	\arrow[""{name=1, anchor=center, inner sep=0}, "{q_1}"', bend right=40, from=1-2, to=1-3]
    	\arrow[""{name=2, anchor=center, inner sep=0}, "p_0", bend left=40, from=1-1, to=1-2]
    	\arrow[""{name=3, anchor=center, inner sep=0}, "q_0", bend left=40, from=1-2, to=1-3]
    	\arrow["s", shorten <=3pt, shorten >=3pt, Rightarrow, from=3, to=1]
    	\arrow["r"', shorten <=3pt, shorten >=3pt, Rightarrow, from=2, to=0]
        \end{tikzcd}\]

        \item Given paths fitting into the following diagram
    	\[\begin{tikzcd}
        x && y && z
        \arrow[""{name=0, anchor=center, inner sep=0}, "{p_1}"', bend right=40, from=1-1, to=1-3]
    	\arrow[""{name=1, anchor=center, inner sep=0}, "{q_1}"', bend right=40, from=1-3, to=1-5]
    	\arrow[""{name=2, anchor=center, inner sep=0}, "{p_0}", bend left=40, from=1-1, to=1-3]
    	\arrow[""{name=3, anchor=center, inner sep=0}, "{q_0}", bend left=40, from=1-3, to=1-5]
    	\arrow[""{name=4, anchor=center, inner sep=0}, from=1-1, to=1-3]
    	\arrow[""{name=5, anchor=center, inner sep=0}, from=1-3, to=1-5]
    	\arrow["{r_0}"', shorten <=2pt, shorten >=2pt, Rightarrow, from=2, to=4]
    	\arrow["{r_1}"', shorten <=2pt, shorten >=2pt, Rightarrow, from=4, to=0]
    	\arrow["{s_0}", shorten <=2pt, shorten >=2pt, Rightarrow, from=3, to=5]
    	\arrow["{s_1}", shorten <=2pt, shorten >=2pt, Rightarrow, from=5, to=1]
        \end{tikzcd}\]
        show we have the \textbf{interchange law}:
        \[ (s_1 \ast r_1) \cdot (s_0 \ast r_0) = (s_1 \cdot s_0) \ast (r_1 \cdot r_0)  \]

        \item Deduce that for $n \geq 2$, the path composition in $\Omega^n(X, x)$ is commutative, and thus that $\pi_n(X, x)$ is abelian.
    \end{enumerate}
    \item (The Fundamental Theorem of Identity Types)
    
    Let $R : X \to X \to \UU$ be a reflexive binary relation. The following are equivalent   \begin{enumerate}
        \item The obvious map\footnote{i.e.\ the map defined by path induction which
          sends $\refl_x$ to the term of type $R(x,x)$ given to us by
          reflexivity of $R$} $x =_A y \to R(x,y)$  is an equivalence.
        \item For any $x :X$, the total space $(y : X) \times (R(x ,y))$ is contractible.
    \end{enumerate}
 \item Recall that $\UU^{\le n} := (X: \UU) \times \isType{n}(X)$ is an $(n+1)$-type. This allows us to define $R : \Trunc{X}_{n+1} \times \Trunc{X}_{n+1} \to \UU^{\le n}$ by the truncation elimination: 
    \[R(\trunc{x}_{n+1} , \trunc{y}_{n+1}) := \Trunc{x = y}_{n}\]
    %

    Apply the fundamental theorem of identity types with the above
    family to deduce that $(\trunc{x}_{n+1} =_{\Trunc{X}_{n+1}} \trunc{y}_{n+1}) \simeq \Trunc{x = y}_n$
    
  \item Show that when $X$ is connected, that $\pi_n(X, x_0) \cong \pi_n(X, x_1)$ are merely isomorphic for any $x_0, x_1 : X$, that is, construct an element of the type
    \[
      (x_0, x_1: X) \to \Trunc{\pi_n(X,x_0) \simeq \pi_n(X,x_1)}.
    \]

    \item We define the observational equality type family of the natural numbers $E : \NN \to \NN \to \UU$  by
        \begin{align*}
            E(0,0) &= 1 \\
            E(\mathrm{succ}(n), 0) &= 0 \\
            E(0, \mathrm{succ}(n)) &= 0 \\
            E(\mathrm{succ}(n), \mathrm{succ}(m)) &= E(n,m)
        \end{align*}
        \begin{enumerate}
            \item Define a natural map $n = m \to E(n,m)$ for all $n, m : \NN$.
            \item Use the fundamental theorem of identity types to show this natural map is an equivalence.
        \end{enumerate}

  \item Show that a type $X$ is $n+1$-truncated if and only if the diagonal map of $X$,
    \[\delta_X : X \to X\times X,\quad x\mapsto (x,x),\]
    is $n$-truncated.


  \item Show that $\Trunc X_0$ satisfies the universal property of the set
    quotient of $X$ modulo the equivalence relation $\sim$, where $x\sim x'$ if
    and only if $\Trunc{x=_X x'}$.

  %% \item Given two functions $f : X_1 \to Y$ and $g : X_2 \to Y$, we
  %%   can define their pullback $X_1 \times_Y X_2 := (x_1 : X_1) \times
  %%   (x_2 : X_2) \times (f(x_1) = g(x_2))$. Letting $\UU/Y : = (X :
  %%   \UU) \times (X \to Y)$, this can be viewed as a product $\times :
  %%   \UU/Y \times \UU/Y \to \UU/Y$ defined by $(X_1,f) \times (X_2,g)
  %%   := (X_1 \times_Y X_2 , (x_1,x_2,p) \mapsto f(x_1))$
  %%   \begin{enumerate}
  %%   \item Define the functorial action of $\times$. For this question, you should is a triple $(\alpha,\beta,\gamma)$.
  %%   \item An Define the functorial action
  %%   \end{enumerate}


% \begin{enumerate}
%   \item The obvious way of defining the type of arrows over the universe $(x:X)$
% \end{enumerate}
% This means, in particular, that we have an automorpism $(x:X) \times (X \to Y) \simeq (x : X) \times(X \to Y)$ defined by
% \[(X \xrightarrow{f} Y) \mapsto ((y : Y) \times (\fib_f(y)) \xrightarrow{\fst} Y).\]
% As a result, anytime we wish to prove a property ranging over the the slice $\UU/Y := $
\end{enumerate}
\section{Solutions}
\begin{enumerate}
    \item Define $f : \| \,\|A\|_n\|_k \to \|A\|_k$ to be $f (||a|_n|_k)= |a|_k$ and $g: \|A\|_k \to \|\, \|A\|_n\|_k$ to be $g(|a|_k)=||a|_n|_k$. Both functions are defined using the universal property of truncations, and the fact that $\|A\|_k$ is $n$-truncated because $k \leq n$. It's then easy to see that $f$ and $g$ define a quasi-eqivalent.

    
    \item First assume $X$ is an $n$-connected type. Then $\|X\|_n$ is contractible, and hence for every $n$-type $Y$, we apply the universal property of $n$-truncation and get:
    \[
    (X \to Y) \simeq (\|X\|_n \to Y) \simeq Y
    \]
    Therefore, we can identify $\delta$ with $id_Y$, and so it is an equivalence.

    Conversely, let $\delta_Y : Y \to (X \to Y)$ defines an equivalence for all $n$-types $Y$. So in particular it is an equivalence for $Y:= \|X\|_n$. Applying this on $|\mbox{-}|_n : X \to \|X\|_n$, we get:
    \[
     \fib_{\delta_{\|X\|_n}}|\mbox{-}|_n
    \]
    is contractible. Unfolding this, we get that the type
    \[
    (z : \|X\|_n) \times (\mathsf{const}_z = |\mbox{-}|_n)
    \]
    is contractible. By function extensionality, we get that
    \[
    (z : \|X\|_n) \times ((x : X) \to |x|_n = z)
    \]
    is contractible. Now since identities of $\|X\|_n$ are $(n-1)$-truncated, and so $n$-truncated, we can apply universal property of truncation to get that $\|X\|_n$ is contractible.
    \item First, show that the first projection map from $(z : \|A\|_n) \times \| \fib_{|\mbox{-}|_n}z\|_n$ to $\|A\|_n$ is an equivalence, using that the domain is an $n$-type. Then apply the following lemma:
    \begin{lemma}
        For any type $A$ and a family of types $B : A \to \UU$, the total space $(x: A) \times B(x)$ is contractible if and only of for each $x:A$, $B(x)$ is a contractible type.
    \end{lemma}
    \item The map $i$ is clearly $n$-truncated as its fibers are $n$-truncation and therefore $n$-types. We can identify $X$ with the total space of fibers $(y:Y) \to \fib_f(y)$. Then $p$ is $n$-connected by observing that the map between total spaces is $n$-connected if maps between fibers are, and the later is the result of the previous exercise.
    \item
    \item 
    \item 
    \item Since $X$ is a connected type, we have a map $f$ assigns to each two points $x_1,x_2 : X$ a path $f(x_1, x_2) :|x_1|_0 =_{\|X\|_0} |x_2|_0$. On the other hand, we know that any path $p: x_1 = x_2$ implies the goal, and since  the goal is a proposition, it is sufficient to find a proof for $\| x_1 = x_2\|$, which can be found by applying the following lemma.
    \begin{lemma}
For any type $X$ and any $x_1,x_2:X$, there is an equivalence
\[
\bigl(\,|x_1|_0 =_{\|X\|_0} |x_2|_0\,\bigr)\;\simeq\;\|\,x_1 =_X x_2\,\|.
\]
\end{lemma}

\begin{proof}
Since $\|X\|_0$ is a set, the type
\[
|x_1|_0 =_{\|X\|_0} |x_2|_0
\]
is a proposition; and $\|x_1=x_2\|$ is also a proposition. Hence it suffices to give maps in both directions.

For
\[
\|x_1=x_2\| \to \bigl(|x_1|_0 = |x_2|_0\bigr),
\]
use propositional truncation recursion and send a path
\[
p:x_1=x_2
\]
to
\[
\ap_{|-|_0}(p): |x_1|_0 = |x_2|_0.
\]

For the converse, define the proposition-valued family
\[
P:\|X\|_0 \to \mathcal U,
\qquad
P(z):\equiv \|\,x_1 = z\,\|,
\]
where $P$ is induced from the family $y \mapsto \|x_1=y\|$ on $X$ by set-truncation recursion. Then $P(|x_1|_0)$ is inhabited by $|\refl_{x_1}|$. Thus, given
\[
q:|x_1|_0 = |x_2|_0,
\]
transporting along $q$ yields an inhabitant of
\[
P(|x_2|_0)\equiv \|x_1=x_2\|.
\]

Therefore the two proposition-valued types are equivalent. Note that one can generalize this result to $n$-truncation (see the HoTT book, Theorem 7.3.12).
\end{proof}
    \item
    \item For each $(x_1 , x_2) : X \times X$, we have:
    $$\fib_{\delta_X} (x_1,x_2) = (x:X) \times ((x,x) = (x_1,x_2)) = (x:X) \times (x=x_1 \times x=x_2) = x_1 = x_2  ,$$
    and we know that $X$ is $(n+1)$-truncated if and only if its identity types are $n$-truncated.
    \item The relation
\[
x\sim x' \;:\!\!\equiv\; \Trunc{x=_X x'}
\]
is an equivalence relation: reflexivity is given by $|\refl_x|$, while symmetry and
transitivity follow by truncation elimination into propositions.

Now let $A$ be a set. To show that $\Trunc X_0$ is the set quotient of $X$ by $\sim$,
we must show that precomposition with the truncation map
\[
|-|_0 : X \to \Trunc X_0
\]
induces an equivalence
\[
(\Trunc X_0 \to A)\;\simeq\;\Bigl\{\,f:X\to A \mid \forall x,x':X,\; x\sim x' \to f(x)=f(x')\,\Bigr\}.
\]

Given $g:\Trunc X_0\to A$, define $f:X\to A$ by $f(x):=g(|x|_0)$. Since $A$ is a set, the type
$f(x)=f(x')$ is a proposition. Hence by truncation elimination it suffices to
consider $p:x=x'$, in which case
\[
f(x)=g(|x|_0)=g(|x'|_0)=f(x')
\]
by applying $g$ to $\ap_{|-|_0}(p)$.

Conversely, suppose $f:X\to A$ respects $\sim$. Since $A$ is a set, set-truncation
recursion yields a map
\[
\bar f:\Trunc X_0\to A,
\qquad
\bar f(|x|_0):=f(x),
\]
and this is well-defined exactly because $f$ respects $\sim$.

These two constructions are inverse to each other, by the computation rule for
set truncation and by set-truncation induction. Therefore $\Trunc X_0$ satisfies
the universal property of the set quotient of $X$ modulo $\sim$.
    
    

    
\end{enumerate}
\chapter{A menagerie of higher inductive types}


\section{Exercises}
\begin{enumerate}
\item \label{ex-pullpushdual} (Pushout-Pullback Duality) Consider a commutating square:

\[
    \begin{tikzcd}
      A \ar[d,"f"']\ar[r,"g"] & C\ar[d,"{j}"] \\
      B\ar[r,"{i}"'] & X
    \end{tikzcd}
  \]

with $H : i \circ f \sim j \circ g$. Then the following are equivalent:
\begin{enumerate}[label=(\roman*)]
    \item The square is a pushout square.
    \item the square
    \[
    \begin{tikzcd}
      T^X \ar[d,"- \circ i"']\ar[r,"- \circ j"] & T^C\ar[d,"- \circ g"] \\
      T^B\ar[r,"- \circ f"'] & T^A
    \end{tikzcd}
  \]
is a pullback square, for every type $T$.
\end{enumerate}
\item Show that the two definitions of $S^1$ are equivalent, in two ways:
\begin{enumerate}[label=(\roman*)]
    \item Show the
  HIT for $S^1$ is the same thing as $\Sigma 2$ by constructing an equivalence between them.
  \item  By applying Exercise 2.\ref{ex-pullpushdual}.
\end{enumerate}

%% \item Recall the notion of quasi-isomorphism from \cref{rmk:qinv}. Show that $(S^1 \simeq S^1)$

\item Show that evaluation at the loop gives pointed equivalence
  \[
    (S^1 \pto X) \pto \Omega X
  \]
  for any pointed type $X$.

  Conclude, using the suspension--loop adjunction, that we have a natural equivalence
  \[
    (S^n \pto X) \pto \Omega^n X
  \]
  for all $n:\NN$.

\item Show that a type $X$ is a proposition iff $\inl : X \to X \ast X$ is an equivalence.

\item Show that $A\vee B \simeq A*B$ for propositions $A,B$, where
  $A\vee B := \Trunc{A+B}$.  (Beware the notational overload between the wedge
  sum and disjunction: It's not too bad in practice, because the only pointed
  proposition is the unit type!)

\item \begin{enumerate}[label=(\roman*)]
    \item Consider a span $A \xleftarrow[]{f}X \xrightarrow[]{g}B$. Show that the following square is pushout:
    \[
    \begin{tikzcd}
      X + X \ar[d,"f+g"']\ar[r,"\text{[id,id]}"]\pocorner & X\ar[d,"{inr \circ g}"] \\
      A+B\ar[r,"\text{[inl,inr]}"'] & A \amalg_X B
    \end{tikzcd}
  \]
  \item Apply this to show that for every type X we get $\Sigma X \simeq \mathsf{bool}*X$.
\end{enumerate}. (Harder)
  Show that $S^1 \wedge X \simeq \Sigma X$

\item (Smash Product - Pointed Map Space Adjunction) 
\begin{enumerate}[label=(\roman*)]
    \item Prove that for each pointed type $B$, the fucnction $ (-) \land B$ is the left adjoint to the function $B \to_* (-) $ ,i.e. for every pointed types $A$ and $C$, there is a pointed equivalence:
    $$
    A \to_* ( B \to_* C) \simeq_* (A \land B) \to_* C
    $$
    \item Use this to prove that  $S^1 \land X \simeq \Sigma X$ for each type $X$.
\end{enumerate}

\item Show that the smash product is equivalently the pushout:
  \[
    \begin{tikzcd}
      X + Y \ar[d]\ar[r] & X \times Y\ar[d]\\
      1 + 1 \ar[r] & X \land Y
    \end{tikzcd}
  \]
\item \begin{enumerate} [label=(\roman*)]
    \item For all pointed types $X$, $Y$, and $A$, prove there is a pointed equivalence:
    $$
    X * Y \to_* A \,\simeq_* \, X \to (Y \to_* \Omega A) 
    $$
    \item For pointed types $X,Y$ prove that
  \[X * Y \simeq_* \Sigma (X\wedge Y).\]
\end{enumerate} 

% \item The torus, $T^2$, is often drawn as a (filled) square with side identifications
% \begin{equation}\label{eq:torus}
% \begin{tikzcd}[ampersand replacement=\&]
% 	\star \& \star \\
% 	\star \& \star
% 	\arrow["{\ell_1}", from=1-1, to=1-2]
% 	\arrow["{\ell_2}", from=2-1, to=1-1]
% 	\arrow["{\ell_1}"', from=2-1, to=2-2]
% 	\arrow["{\ell_2}"', from=2-2, to=1-2]
% \end{tikzcd}
% \end{equation}
% In HoTT, we can capture this as HIT with
% \begin{enumerate}
% \item[$\bullet$] A point constructor $\star : T^2$
% \item[$\bullet$] Two paths $\ell_1,\ell_2 : \star = \star$
% \item[$\bullet$] A filler $\mathsf{sq}$ of the square of paths in \eqref{eq:torus}.
% \end{enumerate}
%Recall \cref{def:fillers} for the definition of \emph{fillers}.
% %
% Show this type is equivalent to $S^1 \times S^1$.
% %
  % (\emph{Hint})\enspace You may assume that eliminators compute definitionally
  % on point constructors (as in cubical type theory), otherwise it's quite
  % intricate!

\item We can define the unit interval type $\II$ to be the pushout of the constant span $1 \leftarrow 1 \rightarrow 1$. Working in Martin-Löf type theory \emph{without} univalence (and function extensionality) but with pushouts with strict definitional computation rules, prove that
  \begin{enumerate}[label=(\roman*)]
  \item $ [((x, y) : A \times A) \times (x = y)] \simeq (\II \to A)$
  \item Use the above equivalence (or, if you prefer, a variation thereof) to derive function extensionality.
  \end{enumerate}
  
\item \begin{enumerate}[label=(\roman*)]
    \item Show that if the square
    \[
    \begin{tikzcd}
      A \ar[d,"f"']\ar[r,"g"] & C\ar[d,"{j}"] \\
      B\ar[r,"{i}"'] & X
    \end{tikzcd}
  \]
is a pullback, then so is
\[
    \begin{tikzcd}
      \Omega A \ar[d,"\Omega f"']\ar[r,"\Omega g"] & \Omega C\ar[d,"{\Omega j}"] \\
      \Omega B\ar[r,"{\Omega i}"'] & \Omega X
    \end{tikzcd}
  \]
  \item For a map $A \xrightarrow[]{f}B$ show that $\Omega \, (\fib \,f) \simeq \fib \, (\Omega f)$
\end{enumerate}
  \item Show that the square
    \[
    \begin{tikzcd}
      S^1 \ar[d,""']\ar[r,""] & 1\ar[d,"{}"] \\
      S^1 \times S^1 \ar[r,"{}"'] & S^2 \lor S^1
    \end{tikzcd}
  \]
  is a pushout.
\end{enumerate}
\section{Solutions}
\begin{enumerate}
\item 
\item 
\item 
\item 
\item 
\item 
\item 
\item 
\item 
\item 
Write the interval as the pushout of the constant span $1 \leftarrow 1 \rightarrow 1$, with points 
$0,1:\II $
and path
$
\mathsf{seg}:0=1
$.

\begin{enumerate}[label=(\roman*)]
\item Define
\[
\Phi:\Bigl(\sum_{x,y:A}(x=y)\Bigr)\to (\II\to A)
\]
by interval recursion:
\[
\Phi(x,y,p)\;:=\;\mathsf{rec}_{\II}(x,y,p).
\]
Thus, by the strict $\beta$-rules, we get:
\[
\Phi(x,y,p)(0)\equiv x,\qquad
\Phi(x,y,p)(1)\equiv y,\qquad
\ap_{\Phi(x,y,p)}(\mathsf{seg})\equiv p.
\]

Conversely, define
\[
\Psi:(\II\to A)\to \sum_{x,y:A}(x=y)
\]
by
\[
\Psi(h)\;:=\;\bigl(h(0),\,h(1),\,\ap_h(\mathsf{seg})\bigr).
\]

Then
\[
\Psi(\Phi(x,y,p))
\equiv
(x,y,p)
\]
judgmentally, by the strict $\beta$-rules above. For the other composite, if $h:\II\to A$, then
\[
\Phi(\Psi(h))
=
\Phi\bigl(h(0),h(1),\ap_h(\mathsf{seg})\bigr)
\]
is the map out of $\II$ with the same values on $0,1$ and the same action on $\mathsf{seg}$ as $h$; hence, by the $\eta$-rule of the pushout $\II$,
\[
\Phi(\Psi(h)) = h.
\]
Therefore
\[
\Bigl(\sum_{x,y:A}(x=y)\Bigr)\simeq (\II\to A).
\]
Note that we do not actually need strict $\beta$-rule for this part.

\item First we derive non-dependent function extensionality. Let $f,g:A\to B$ and assume
\[
H:\prod_{a:A}(f(a)=g(a)).
\]
By part (i), for each $a:A$ we obtain a map
\[
\widetilde H(a):\II\to B
\]
such that
\[
\widetilde H(a)(0)\equiv f(a),\qquad
\widetilde H(a)(1)\equiv g(a),\qquad
\ap_{\widetilde H(a)}(\mathsf{seg})\equiv H(a).
\]
Now curry these maps:
\[
K:\II\to (A\to B),\qquad
K(i)(a):=\widetilde H(a)(i).
\]
Then, using the strict computation rules from part (i) and the judgmental $\eta$-rule for functions,
\[
K(0)\equiv \lambda a.\,\widetilde H(a)(0)\equiv \lambda a.\,f(a)\equiv f,
\]
and similarly
\[
K(1)\equiv g.
\]
Hence
\[
\ap_K(\mathsf{seg}) : f=g.
\]
So ordinary function extensionality holds.

To obtain dependent function extensionality, let $B:A\to \mathcal U$, let
\[
f,g:\prod_{a:A}B(a),
\qquad
H:\prod_{a:A}(f(a)=g(a)).
\]
Define
\[
f',g':A\to \sum_{a:A}B(a),
\qquad
f'(a):=(a,f(a)),\quad g'(a):=(a,g(a)).
\]
For each $a:A$, using $H(a)$ we get
\[
f'(a)=g'(a)
\]
in the $\Sigma$-type, namely by the path $(\refl_a,H(a))$. By ordinary function extensionality, this yields
\[
p:f'=g'.
\]
Now apply $\ap$ to the map
\[
s:\Bigl(A\to \sum_{a:A}B(a)\Bigr)\to \prod_{a:A}B(a),
\qquad
s(k)(a):= \pi_2(k(a)).
\]
Since $s(f')=f$ and $s(g')=g$, we obtain
\[
\ap_s(p):f=g.
\]
Thus full function extensionality follows.
\color{red}
It seems that we can prove (ii) without assuming $\eta$-rule for $\II$, right? Since we only need the right to left map from (i), and not the whole equivalence.
\color{black}
\end{enumerate}
\item 
\begin{enumerate}[label=(\roman*)]
    \item Assume the given square is a pullback. Then
    $A \simeq B \times_X C$.
    Since for any pointed type $Y$ we have
    \[
    \Omega Y \simeq 1 \times_Y 1,
    \]
    and pullbacks commute with pullbacks (by 3-3 lemma for pullbacks), it follows that
    \[
    \Omega A
    \;\simeq\;
    \Omega(B \times_X C)
    \;\simeq\;
    (1 \times_B 1)\times_{(1 \times_X 1)}(1 \times_C 1)
    \;\simeq\;
    \Omega B \times_{\Omega X} \Omega C.
    \]

    \item The homotopy fiber of $f$ is the pullback
    \[
    \fib(f) \simeq 1 \times_B A.
    \]
    Apply (i) to this pullback square. 
\end{enumerate}

    \item Using the equation $S^1 \land X \simeq \Sigma X$ we can identify $S^2$ with $S^1 \land S^1$, which is pushout of $S^1 \lor S^1 \to S^1 \times S^1$ along the unit map. Now we have:
      \[
    \begin{tikzcd}
      S^1 \lor S^1 \ar[d,""']\ar[r,""] & 1\ar[d,"{}"] \ar [r] & S^1 \ar[d] \\
      S^1 \times S^1 \ar[r,"{}"'] & S^2 \ar [r] & S^2 \lor S^1
    \end{tikzcd}
    \]
    Since both squares are pushouts, the pasting lemma for pushout says that the outer rectangle is a pushout. We get the goal by observing that the upper square in the following diagram is a pushout:
      \[
    \begin{tikzcd}
      S^1 \ar[d,""']\ar[r,""] & 1\ar[d,"{}"] \\
      S^1 \lor S^1 \ar[r,"{}"'] \ar[d] & S^1 \ar[d] \\ 
      S^1 \times S^1 \ar [r] & S^2 \lor S^1
    \end{tikzcd}
    \]
\end{enumerate}

\chapter[The Hopf fibration and the LES]{The Hopf fibration and the long exact sequence}



\section{Exercises}
\begin{enumerate}
\item Construct the product on the circle directly by double circle induction.
\item Give another proof of~\cref{thm:S1-InfCyc} by showing that
  $\ap_{\varphi} : (\ast=\ast) \to (\varphi(\ast)=\varphi(\ast))$ is an
  equivalence.
\item Construct a retraction $\rho : \Omega S^2 \to S^1$ of $\eta : S^1 \to \Omega S^2$.
  (\emph{Hint}) Use the Hopf fibration.
\item We can define a ``tensor product'' of $\ZZ$-torsors by setting
  \[
    (X,t) \otimes (Y,u) := (X \times Y/\sim, s)
  \]
  using the set quotient by the equivalence relation generated by
  $(t(x),y) \sim (x,u(y))$, and setting $s[(x,y)] := [t(x),y)]$.  Check that
  this operation is well defined and gives an H-space structure on $\Tors_\ZZ$
  with the base point $(\ZZ,\suc)$ as neutral element.
\item Investigate some other components of $\SetEndo$, in particular those
  containing $(\Fin n,\suc)$, where $\Fin n=(k:\NN)\times(k<n)$ and the
  successor operation is taken modulo $n$.
  If this component is denoted $\Cyc_n$ (for $n$-cycles), what are the maps $\Cyc_n \to \Cyc_m$?
\end{enumerate}

\chapter[Freudenthal and Degrees]{The Freudenthal Suspension Theorem and Degrees}
\label{ch:pushouts}

\section{Exercises}

\begin{enumerate}
\item Deduce the dual Blakers--Massey~\cref{thm:dual-BM} from the Join
  Connectivity~\cref{lem:join-connectivity} by identifying the fibers of the
  cogap map with the joins of the fibers of legs in the cospan.

\item Deduce the Wedge Connectivity Lemma from the dual Blakers--Massey~\cref{thm:dual-BM},
  i.e., consider the pullback square, for pointed types $X,Y$,
  \[
    \begin{tikzcd}
      1 \ar[d]\ar[r]\pbcorner & Y \ar[d] \\
      X \ar[r] & X\times Y.
    \end{tikzcd}
  \]
  Show that if $X$ is $n$-connected and $Y$ is $m$-connected, then the wedge
  inclusion map $X \vee Y \to X \times Y$ is $(n+m)$-connected.

\item Show that the counit $\varepsilon : \Sigma\Omega X \to X$ of the loop--suspension adjunction is $2n$-connected if $X$ is $n$-connected.

\item Show that the pushout of an embedding is an embedding, i.e., if $g$ is $(-1)$-truncated, then so is $\inl$ in the pushout square:
  \[
    \begin{tikzcd}
      R \ar[d,"f"']\ar[r,"g"]\pocorner & B\ar[d,"{\inr}"] \\
      A\ar[r,"{\inl}"'] & Q
    \end{tikzcd}
  \]
\end{enumerate}


\raggedright
%\printbibliography

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
